\ttitle

### 2014-01-06 Chapters/Chapter5/quality_metrics.tex
Changed Benefit to Delay. WIP chapter 6

% !TEX root = ../../main.tex
\section{Quality metrics}\label{sec:artificial_data_quality_metrics}
For an objective measure of quality of the proposed method applied to the four aforementioned data sets, we employ two metrics, inspired by and slighty altered from Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
The first metric is the $\operatorname*{average \_ benefit}$, for which we first need to define the notion of $\operatorname*{benefit}$ for each change:
\begin{equation}\label{eq:benefit}
  \operatorname*{benefit}(y) = 1 - |t_y - t_y^*|,
\end{equation}
where $t_y$ is the time of the closest discovered change point and $t_y^*$ is the true change time.
We can then define the $\operatorname*{average \_ benefit}$ over all the discovered change points $Y$ as:
\begin{equation}\label{eq:average_benefit}
  \operatorname*{average \_ benefit}(Y) = \frac{\sum_{y=1}^Y \operatorname*{benefit}(y)}{Y}.
\end{equation}
Besides the benefit, which rewards discovered change points close to the actual change points, we need to penalize discovered change points which do not relate to a true change point.
For this we use the \acrlong{far}, expressed as
\begin{equation}\label{eq:false_alarm_rate}
  \operatorname*{far}(Y) = \frac{Y_{\operatorname*{false}}}{Y}.
\end{equation}

For a high quality method the $\operatorname*{average\_benefit}$ should be high and the $\operatorname*{far}(Y)$ should be low.
In our results we will also list the parameters used, since these need to be user-defined and are part of the (manual) optimization process.
The parameters are:
\begin{itemize}
  \item \textbf{Window length}: The window length that is processed by the \gls{svdd} algorithm. This is the number of data points that are continuously modeled in the shape in a hypersphere.
  \item \textbf{Sigma of \gls{rbf}}: The value of $\sigma$ for the \gls{rbf} kernel. The effect of this parameter is discussed in \Cref{subsec:svm_model_parameters}.
  \item \textbf{High threshold}: The $th_\text{high}$ threshold value for the \gls{rt} method described in \Cref{subsec:ratio_thresholding}. A higher value indicates a more precise and robust setting.
  \item \textbf{Low threshold}: The $th_\text{low}$ threshold value for the \gls{rt} method described in \Cref{subsec:ratio_thresholding}. A lower value indicates a more precise and robust setting.
  \item \textbf{Closeness}: In the post-processing phase the detected change points which are within a certain \emph{closeness} range will be merged together. This parameter indicates the minimal number of data points that change points must differ in order to be considered separate detections.
\end{itemize}
For all the sets we have used a value of $C = 0.1$, which indicated that $10\%$ of the data points are used as outliers.
This is a normal setting for the \gls{svdd} method.

>>>>

% !TEX root = ../../main.tex
\section{Quality metrics}\label{sec:artificial_data_quality_metrics}
For an objective measure of quality of the proposed method applied to the four aforementioned data sets, we employ two metrics, inspired by and slighty altered from Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
The first metric is the $\operatorname*{average \_ delay}$, for which we first need to define the notion of $\operatorname*{delay}$ for each change:
\begin{equation}\label{eq:delay}
  \operatorname*{delay}(y) = |t_y - t_y^*|,
\end{equation}
where $t_y$ is the time of the closest discovered change point and $t_y^*$ is the true change time.
We can then define the $\operatorname*{average \_ delay}$ over all the discovered change points $Y$ as:
\begin{equation}\label{eq:average_delay}
  \operatorname*{average \_ delay}(Y) = \frac{\sum_{y=1}^Y \operatorname*{delay}(y)}{Y}.
\end{equation}
Besides the delay, which should be low to indicate discovered change points near the actual change point, we need to penalize discovered change points which do not relate to a true change point.
For this we use the \acrlong{far}, expressed as
\begin{equation}\label{eq:false_alarm_rate}
  \operatorname*{far}(Y) = \frac{Y_{\operatorname*{false}}}{Y}.
\end{equation}

For a high quality method the $\operatorname*{average\_delay}$ and the $\operatorname*{far}(Y)$ should both be low.
In our results we will also list the parameters used, since these need to be user-defined and are part of the (manual) optimization process.
The parameters are:
\begin{itemize}
  \item \textbf{Window length}: The window length that is processed by the \gls{svdd} algorithm. This is the number of data points that are continuously modeled in the shape in a hypersphere.
  \item \textbf{Sigma of \gls{rbf}}: The value of $\sigma$ for the \gls{rbf} kernel. The effect of this parameter is discussed in \Cref{subsec:svm_model_parameters}.
  \item \textbf{High threshold}: The $th_\text{high}$ threshold value for the \gls{rt} method described in \Cref{subsec:ratio_thresholding}. A higher value indicates a more precise and robust setting.
  \item \textbf{Low threshold}: The $th_\text{low}$ threshold value for the \gls{rt} method described in \Cref{subsec:ratio_thresholding}. A lower value indicates a more precise and robust setting.
  \item \textbf{Closeness}: In the post-processing phase the detected change points which are within a certain \emph{closeness} range will be merged together. This parameter indicates the minimal number of data points that change points must differ in order to be considered separate detections.
\end{itemize}
For all the sets we have used a value of $C = 0.1$, which indicated that $10\%$ of the data points are used as outliers.
This is a normal setting for the \gls{svdd} method.

### 2014-01-06 Chapters/Chapter5/results.tex
Changed Benefit to Delay. WIP chapter 6

\begin{table}
  \centering
  \begin{tabulary}{\textwidth}{|l|c|c|c|c|}
    \cline{2-5}
    \multicolumn{1}{l|}{} & Set 1 & Set 2 & Set 3 & Set 4 \\
    \hline
    Window length & 50 & 100 & 50 & 50 \\
    \hline
    Sigma of \gls{rbf} & 13 & 13 & 15 & 13 \\
    \hline
    High threshold & 1.6 & 1.6 & 1.5 & 2.2 \\
    \hline
    Low threshold & 0.1 & 0.1 & 0.5 & 0.1 \\
    \hline
    Closeness & 10 & 10 & 10 & 50 \\
    \hline
    \hline
    $\operatorname*{far}(Y)$ & 0 & 0.1 & 0.1 & 0.1 \\
    \hline
    $\operatorname*{Average\_benefit}$ & 2.27 & 6.18 & 0.64 & 15.36 \\
    \hline
    Standard deviation benefit & 1.19 & 7.94 & 0.92 & 17.70 \\
    \hline
  \end{tabulary}
  \caption[Results artificial data sets]{Parameter settings and results of the artificial data sets.}
  \label{tab:results_artificial}
\end{table}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter5/boxplot_results_artificial_sets.eps}
  \caption[Box plot results artificial data sets]{Box plot of the results for the artificial data sets, indicating the number of data points difference between the actual and detected change points.}
  \label{fig:boxplot_artificial_sets}
\end{figure}

To give a better impression of the spread of the accuracy of the detected change points, a box plot for each data set is displayed in \Cref{fig:boxplot_artificial_sets} and are the results are visualized in \Cref{fig:plots_results_artificial_data_sets}.
With each data set embodying different characteristics, different observations can be drawn from the data sets.
In the references figures, the blue lines represent the radius of the constructed hypersphere.
The vertical red lines are the discovered change points.
For each data set the following is observed:
\begin{enumerate}
  \item \textbf{Fixed increasing mean:} In \Cref{fig:camci_fixed_increasing_mean_thresholds} we see that each change point is almost equally fast and accurate detected.
  This implies that the relative difference between the means of the Gaussian distribution has no effect on the detectability of the algorithm.
  \item \textbf{Reduced increasing mean:} \Cref{fig:takeuchi_reduced_increasing_mean_thresholds} shows that the absolute difference between the means of the Gaussian distribution is a certain influence to the detectability.
  Although all change points are correctly detected (at the beginning of the series are two outliers), the change in radius becomes less significant at every change point.
  \item \textbf{Reduced increasing mean, increasing variance:} In \Cref{fig:camci_takeuchi_alternating_variance_thresholds} we see the same effect, in regard to the mean, as with the previous data set.
  Furthermore, the increasing variance makes change detection harder.
  This is visible around the $9000$\textsuperscript{th} data point and after, where the radius keeps increasing.
  \item \textbf{Alternating variance:} The ability to detect decreases in variance of the Gaussian distribution is displayed in \Cref{fig:camci_reduced_increasing_mean_increasing_variance_thresholds}.
  Although some change points are discovered with a delay almost up to the length of the window processing the data, the decreases are all detected.
  The box plot of this set, number 4 in \Cref{fig:boxplot_artificial_sets}, shows a wider spread for the results.
\end{enumerate}

>>>>

\begin{table}
  \centering
  \begin{tabulary}{\textwidth}{|l|c|c|c|c|}
    \cline{2-5}
    \multicolumn{1}{l|}{} & Set 1 & Set 2 & Set 3 & Set 4 \\
    \hline
    Window length & 50 & 100 & 50 & 50 \\
    \hline
    Sigma of \gls{rbf} & 13 & 13 & 15 & 13 \\
    \hline
    High threshold & 1.6 & 1.6 & 1.5 & 2.2 \\
    \hline
    Low threshold & 0.1 & 0.1 & 0.5 & 0.1 \\
    \hline
    Closeness & 10 & 10 & 10 & 50 \\
    \hline
    \hline
    $\operatorname*{far}(Y)$ & 0 & 0.1 & 0.1 & 0.1 \\
    \hline
    $\operatorname*{Average\_delay}$ & 2.27 & 6.18 & 0.64 & 15.36 \\
    \hline
    Standard deviation Delay & 1.19 & 7.94 & 0.92 & 17.70 \\
    \hline
  \end{tabulary}
  \caption[Results artificial data sets]{Parameter settings and results of the artificial data sets.}
  \label{tab:results_artificial}
\end{table}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter5/boxplot_results_artificial_sets.eps}
  \caption[Box plot results artificial data sets]{Box plot of the results for the artificial data sets, indicating the number of data points between the actual and closest detected change points. A lower and more compact box plot is better.}
  \label{fig:boxplot_artificial_sets}
\end{figure}

To give a better impression of the spread of the accuracy of the detected change points, a box plot for each data set is displayed in \Cref{fig:boxplot_artificial_sets} and are the results are visualized in \Cref{fig:plots_results_artificial_data_sets}.
This last figure shows the number of data points between each actual and closest detected change point.
Using the box plot notation, we can visualize not only the average but also the spread of the delay.
With each data set embodying different characteristics, different observations can be drawn from the data sets.
In the references figures, the blue lines represent the radius of the constructed hypersphere.
The vertical red lines are the discovered change points.
For each data set the following is observed:
\begin{enumerate}
  \item \textbf{Fixed increasing mean:} In \Cref{fig:camci_fixed_increasing_mean_thresholds} we see that each change point is almost equally fast and accurate detected.
  This implies that the relative difference between the means of the Gaussian distribution has no effect on the detectability of the algorithm.
  \item \textbf{Reduced increasing mean:} \Cref{fig:takeuchi_reduced_increasing_mean_thresholds} shows that the absolute difference between the means of the Gaussian distribution is a certain influence to the detectability.
  Although all change points are correctly detected (at the beginning of the series are two outliers), the change in radius becomes less significant at every change point.
  \item \textbf{Reduced increasing mean, increasing variance:} In \Cref{fig:camci_takeuchi_alternating_variance_thresholds} we see the same effect, in regard to the mean, as with the previous data set.
  Furthermore, the increasing variance makes change detection harder.
  This is visible around the $9000$\textsuperscript{th} data point and after, where the radius keeps increasing.
  \item \textbf{Alternating variance:} The ability to detect decreases in variance of the Gaussian distribution is displayed in \Cref{fig:camci_reduced_increasing_mean_increasing_variance_thresholds}.
  Although some change points are discovered with a delay almost up to the length of the window processing the data, the decreases are all detected.
  The box plot of this set, number 4 in \Cref{fig:boxplot_artificial_sets}, shows a wider spread for the results.
\end{enumerate}

### 2014-01-06 Chapters/Chapter6.tex
WIP: Finalizing chapter 3
Changed Benefit to Delay. WIP chapter 6

%----------------------------------------------------------------------------------------
\section{Outline}
\emph{Not intented for the reader.}
\begin{itemize}
  \item Apply proposed method to real-world applications, such as
    \begin{itemize}
      \item Daily life activity recognition (as the original context of this thesis is)
      \item PowerHouse sensor data
      \item Stock data?
    \end{itemize}
  \item Relate back to filter vs. wrapper methods - give results with different methods?
\end{itemize}

This chapter discusses the real-world data sets used for this research.
In \Cref{sec:data_sets} we discuss earlier used data sets and argue that these are inappropriate for our research.
It is followed by an overview of the setup we have created to record continuous data sets from an uncontrolled environment, both in- and outdoor.
The activities performed are discussed in \Cref{sec:data_gathering}.
As discussed in \Cref{Chapter4}, the algorithm is subject to user parameters.
In \Cref{sec:real_world_parameters} the effect and ranges of the parameters on the real world data sets are discussed.
Finally, the performance of the algorithm is discussed in \Cref{sec:real_world_performance}.

\input{./Chapters/Chapter6/data_sets}
\input{./Chapters/Chapter6/data_gathering}
\input{./Chapters/Chapter6/parameters}
\input{./Chapters/Chapter6/performance}

>>>>

%----------------------------------------------------------------------------------------
\section{Outline}

This chapter discusses the real-world data sets used for this research.
In \Cref{sec:data_sets} we discuss earlier used data sets and argue that these are inappropriate for our research.
It is followed by an overview of the setup we have created to record continuous data sets from an uncontrolled environment, both in- and outdoor.
The activities performed are discussed in \Cref{sec:data_gathering}.
As discussed in \Cref{Chapter4}, the algorithm is subject to user parameters.
% In \Cref{sec:real_world_parameters} the effect and ranges of the parameters on the real world data sets are discussed.
Finally, the results of the algorithm is discussed in \Cref{sec:real_world_performance}.

\TODO{Check for final contents and ordering.}

\input{./Chapters/Chapter6/data_sets}
\input{./Chapters/Chapter6/data_gathering}
% \input{./Chapters/Chapter6/parameters}

% \TODO{Add short section on quality metrics}

\input{./Chapters/Chapter6/results}

### 2014-01-06 Chapters/Chapter6/data_gathering.tex
Changed Benefit to Delay. WIP chapter 6

% !TEX root = ../../main.tex
\section{Data Gathering}\label{sec:data_gathering}
For our data sets of real world activities we have sampled both in- and outdoor activities.
The subjects performing the activities are three males, in an age around 25.
We have used two series of recordings.
The first series are various outdoor recordings with segments of walking, running and standing.
Some movement activities are performed in a straight line, whilst others are performed around corners of in a wide circular direction.
The second series consist of mainly walking up and down the stairs, with small vertical parts between the segments.
The stairs are all in a circular shape.

\subsection{Outdoor straight lines}\label{subsec:outdoor_straight}
The first series of outdoor performed activities consist of walking and running segments in a straight line.
Halfway the recording is a $180^{\circ}$ turn.
The activities consist of standing still, walking and running (both jogging and sprinting).
The plots of \Cref{fig:data_gathering_run_1_rot,fig:data_gathering_run_1_acc,fig:data_gathering_run_1_mag} shows the recordings of Subject $1$.
For Subject $2$ the plots are visible in \Cref{fig:data_gathering_run_2_rot,fig:data_gathering_run_2_acc,fig:data_gathering_run_2_mag}.
The recorded metrics are from the accelerometer, rotational and magnetometer sensors.
The solid black lines indicate manually annoted change points, the dashes lines indicate the closest discovered change point.
A still from the video recording, showing Subject $2$ performing a (sprinting) run segment, is in \Cref{fig:stills_subject_2_sprinting}.

\subsection{Outdoor straight lines results}
\TODO{For each data set: discovered timepoint with acc, mag, rot, and all combined. Use closest change point}

\begin{center}\begin{table}
  \begin{tabulary}{\textwidth}{|l|c|c|}
    \hline
    \multicolumn{3}{|c|}{Subj 1, run 1} \\
    \hline \hline
    Act. & Ann. & Disc. \\
    \hline
    Still & 0 & *0* \\
    \hline
    Shake & 3.5 & *3.5* \\
    \hline
    Still & 4.2 & *4.2* \\
    \hline
    In pocket & 5.5 & *5.5* \\
    \hline
    Walk & 7 & *7* \\
    \hline
    Run & 10 & *10* \\
    \hline
    Walk & 14 & *14* \\
    \hline
    Run & 17.2 & *17.2* \\
    \hline
    Walk & 22.1 & *22.1* \\
    \hline
    Turn CCW & 23.5 & *23.5* \\
    \hline
    Walk & 24.2 & *24.2* \\
    \hline
    Run (sprint) & 28 & *28* \\
    \hline
    Run & 31 & *31* \\
    \hline
    Walk & 33 & *33* \\
    \hline
    Run & 35.5 & *35.5* \\
    \hline
    Walk & 39 & *39* \\
    \hline
    Out pocket & 41 & *41* \\
    \hline
    Still & 43 & *43* \\
    \hline
    \hline
    \multicolumn{2}{|l|}{Ratio Total CPs} & 3.2 \\
    \hline
    \multicolumn{2}{|l|}{Sum of differences} & 12.341 \\
    \hline
  \end{tabulary}
  \caption[Performed activities subject 1 run 1]{The outdoor series performed activities by subject 1, run 1.}
  \label{tab:outdoor_series}
\end{table}\end{center}

\TODO{Replace all times in the stars with the closest discovered change point}

@@

%---- Run 3: Stairs remco
\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_rot_annotated.eps}
  \caption[R3: rotation]{Run 3: Indoor stairs Marc, rotation}
  \label{fig:data_gathering_run_3_rot}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_mag_annotated.eps}
  \caption[R3: mag]{Run 3: Indoor stairs Marc, Mag}
  \label{fig:data_gathering_run_3_mag}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_annotated.eps}
  \caption[R3: accelerometer]{Run 3: Indoor stairs Marc, accelerometer}
  \label{fig:data_gathering_run_3_acc}
\end{figure}

>>>>

% !TEX root = ../../main.tex
\section{Data Gathering}\label{sec:data_gathering}
\TODO{Determine subjects:}
\begin{enumerate}
  \item Roemer?
  \item Jos?
  \item Marc
\end{enumerate}
For our data sets of real world activities we have sampled both in- and outdoor activities.
The subjects performing the activities are three males, of an age around 25.
We have used two series of recordings.
The first series consists of various outdoor recordings with segments of walking, running and standing.
Some activities are performed in a straight line, whilst others are performed around corners or in a wide circular direction.
The second series consists mainly of walking up and down the stairs, with small vertical parts between the segments.
The stairs are all in a circular shape.

\subsection{Outdoor straight lines}\label{subsec:outdoor_straight}
The first series of outdoor performed activities consist of walking and running segments in a straight line.
Halfway the recording is a $180^{\circ}$ turn.
The activities consist of standing still, walking and running (both jogging and sprinting).
The plots of \Cref{fig:plots_subject_1} shows the recordings of Subject $1$.
For Subject $2$ the plots are visible in \Cref{fig:plots_subject_2}.
The recorded metrics originate from the accelerometer, rotational and magnetometer sensors.
The solid black lines indicate manually annoted change points, the dashes lines indicate the closest discovered change point.
A still from the video recording, showing Subject $2$ performing a (sprinting) run segment, is in \Cref{fig:stills_subject_2_sprinting}.

% \begin{center}\begin{table}
%   \begin{tabulary}{\textwidth}{|l|c|c|}
%     \hline
%     \multicolumn{3}{|c|}{Subj 1, run 1} \\
%     \hline \hline
%     Act. & Ann. & Disc. \\
%     \hline
%     Still & 0 & *0* \\
%     \hline
%     Shake & 3.5 & *3.5* \\
%     \hline
%     Still & 4.2 & *4.2* \\
%     \hline
%     In pocket & 5.5 & *5.5* \\
%     \hline
%     Walk & 7 & *7* \\
%     \hline
%     Run & 10 & *10* \\
%     \hline
%     Walk & 14 & *14* \\
%     \hline
%     Run & 17.2 & *17.2* \\
%     \hline
%     Walk & 22.1 & *22.1* \\
%     \hline
%     Turn CCW & 23.5 & *23.5* \\
%     \hline
%     Walk & 24.2 & *24.2* \\
%     \hline
%     Run (sprint) & 28 & *28* \\
%     \hline
%     Run & 31 & *31* \\
%     \hline
%     Walk & 33 & *33* \\
%     \hline
%     Run & 35.5 & *35.5* \\
%     \hline
%     Walk & 39 & *39* \\
%     \hline
%     Out pocket & 41 & *41* \\
%     \hline
%     Still & 43 & *43* \\
%     \hline
%     \hline
%     \multicolumn{2}{|l|}{Ratio Total CPs} & 3.2 \\
%     \hline
%     \multicolumn{2}{|l|}{Sum of differences} & 12.341 \\
%     \hline
%   \end{tabulary}
%   \caption[Performed activities subject 1 run 1]{The outdoor series performed activities by subject 1, run 1.}
%   \label{tab:outdoor_series}
% \end{table}\end{center}

@@

%---- Run 3: Stairs remco
\begin{figure}
  \centering
  \begin{subfigure}[b]{1\textwidth}
    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_rot_annotated.eps}
    \caption{Run 3: Indoor stairs Marc, rotation}
    \label{fig:data_gathering_run_3_rot}
  \end{subfigure} \\

  \begin{subfigure}[b]{1\textwidth}
    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_mag_annotated.eps}
    \caption{Run 3: Indoor stairs Marc, Mag}
    \label{fig:data_gathering_run_3_mag}
  \end{subfigure} \\

  \begin{subfigure}[b]{1\textwidth}
    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_annotated.eps}
    \caption{Run 3: Indoor stairs Marc, accelerometer}
    \label{fig:data_gathering_run_3_acc}
  \end{subfigure}
  \caption[Plots subject 3]{Annotated plots of a full recording performed by Subject~$3$ while walking the stairs. The vertical lines are manually determined change points. The label above each segment indicates the current performed activity.}\label{fig:plots_subject_3}
\end{figure}

### 2014-01-06 Chapters/Chapter6/data_sets.tex
Changed Benefit to Delay. WIP chapter 6

% !TEX root = ../../main.tex
\section{Data Sets}\label{sec:data_sets}
In the previous chapter we have applied our method to the artificial data sets from \cite{camci2010change,takeuchi2006unifying}.
In this chapter we apply the method to real-world data sets.
The data sets are inertial sensor data recorded during human activities, such as walking, running, ascending and descending stairs, and standing still.
To compare our results with other research, we have tried to apply our method to two commonly used data sets.
The first is the data from the \gls{wisdm} lab \cite{kwapisz2011activity}.
As an excerpt from the data in \Cref{fig:wisdm_excerpt} shows, the activities are recorded non-continuous.
Whilst this data set is convenient for activity classification, it does not allow for detection of change points.
The other commonly used data set is the \gls{uci-har} from the Machine Learning Repository and originates from \cite{anguita2012human}.
As shown in an except in \Cref{fig:uci_annotated}, the labeling to the data segments seems to be incorrect.
Furthermore, it is not clear how the data is recorded and whether all activities are performed continuously.

@@

Because of the above described shortcomings of commonly used data sets, we have decided to record activities by ourselves.
The following characterizations were set for our data sets:
\begin{enumerate}
  \item The activities need to be performed and recorded in a continuous matter,
  \item The environment should be natural and uncontrolled, both in- and outdoor,
  \item There should be a clear annotation possible, as objective as possible,
  \item The hardware and software should be widely available.
\end{enumerate}

With these requirements in mind, we created the following setup.
All activities were recorded by one smartphone in the right front pants pocket of the subject.
The phones that were used are the \textsc{HTC Sensation XE} and \textsc{HTC Desire}, running the \textsc{Anroid} smartphone operating system.
During the activities the intertial sensor data is recorded using a free application, \textsc{Sensor Logger}~\cite{sensorlogger}.
This application records the sensor data to \textsc{CSV} files, which are then transformed to \textsc{MATLAB} compatible files.
The constructed algorithm processes the data in an online matter (using only current and historic data).
During the activities, the subject is recorded with a video recorder to enable manual labeling and determining the change points.

Using the video recordings, the data sets are manually annotated, as shown in \Cref{fig:data_gathering_run_1_rot}.
The graph displays from a single recording the rotational-axis values (see \Cref{tab:recorded_metrics} for explanation of the metrics), with manually annotated activities above the data.
The manually determines change points are used to draw quality conclusion over the algorithms performance, by comparing it to the discovered change points.
All used data recordings, including the video material and manual labeling, are made available \cite{vlasveld2013continuous}.
\TODO{Create web page with data sets}
The following section discusses the performed activities during the recordings.

>>>>

% !TEX root = ../../main.tex
\section{Data Sets}\label{sec:data_sets}
In the previous chapter we have applied our method to the artificial data sets from \cite{camci2010change,takeuchi2006unifying}.
In this chapter we apply the method to real-world data sets.
The data sets come from inertial sensor data recorded during human activities, such as walking, running, ascending and descending stairs, and standing still.
To compare our results with other research, we have tried to apply our method to two commonly used data sets.
The first is the data from the \gls{wisdm} lab \cite{kwapisz2011activity}.
As an excerpt from the data in \Cref{fig:wisdm_excerpt} shows, the activities are recorded non-continuous.
Whilst this data set is convenient for activity classification, it does not allow for detection of change points.
Even if the gap between the activities would be removed, the transition segments between activities would still be missing and thus the data set would not reflect continuous recordings.
The other commonly used data set is the \gls{uci-har} from the Machine Learning Repository and originates from \cite{anguita2012human}.
As shown in an except in \Cref{fig:uci_annotated}, the labeling to the data segments seems to be incorrect.
Furthermore, it is not clear how the data is recorded and whether all activities are performed continuously.

@@

Because of the above described shortcomings of commonly used data sets, we have decided to record activities by ourselves.
The following characterization requirements were set for our data sets:
\begin{enumerate}
  \item The activities need to be performed and recorded in a continuous manner,
  \item The environment should be natural and uncontrolled, both in- and outdoor,
  \item There data should be clearly annotated, as objective as possible,
  \item The required hardware (and software) should be widely available.
\end{enumerate}

With these requirements in mind, we have created the following setup.
All activities were recorded by a smartphone in the right front pants pocket of the subject.
The phones that were used are the \textsc{HTC Sensation XE} and \textsc{HTC Desire}, running the \textsc{Anroid} smartphone operating system.
During the activities the intertial sensor data is recorded using a free application, \textsc{Sensor Logger}~\cite{sensorlogger}.
This application records the sensor data to \textsc{CSV} files, which are then transformed to \textsc{MATLAB} compatible files.
The constructed algorithm processes the data in an online manner (using only current and historic data).
During the activities, the subject is recorded with a video camera to enable manual labeling and determination of the change points.

Using the video recordings, the data sets are manually annotated, as shown in the plots of \Cref{fig:plots_subject_1,fig:plots_subject_3}.
The graph displays from a single recording the rotational-axis values (see \Cref{tab:recorded_metrics} for explanation of the metrics), with manually annotated activities above the data.
The manually determined change points are used to draw quality conclusion over the algorithm's performance, by comparing it to the discovered change points.
All used data recordings, including the video material and manual labeling, are made available \cite{vlasveld2013continuous}.
\TODO{Create web page with data sets}
The following section discusses the performed activities during the recordings.
